---
title: "PSYC_201B_COVID_codes"
author: Hyojeong(Jenny) Yoo, Jinah Kim, Juanshu(Sky) Wu, Muqing Fan, Xiangyu(Alaric) Wei, Yuchen(Eva) Liu  
date: "2/28/2021"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Introduction
In this file, we cleaned and explored our following datasets:   
- **JHU covid data (curren)**: https://github.com/CSSEGISandData/COVID-19  
- **Google mobility reports (current)**: https://www.google.com/covid19/mobility/?hl=en  
The JHU COVID-19 date provided us with historical covid progressions. Mobility data provided us relevant information during the pandemic. All datasets are analyzed on state-level.

# Data Cleaning 
Importing packages
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(zoo)
library(imputeTS)
```

## SVI data for county population 
selecting relevant indices (refer to codebook)
```{r, message=FALSE}
svi.raw <- read_csv("./datasets/SVI2018_US_COUNTY.csv")
svi.data <- svi.raw %>%
  select(STATE,
         COUNTY,
         FIPS,
         E_TOTPOP) %>%
  mutate(FIPS = as.numeric(FIPS),
         county = paste(COUNTY, "County"))

```


## JHU COVID-19 data cleaning

loading JHU covid confirmed historical dateset
```{r, message=FALSE}
covid.confirmed.raw <- data.table::fread("./datasets/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")
# glimpse(covid.confirmed.raw)
# cumulative confirmed cases on county level 
```

cleaning JHU covid confirmed historical dateset
### simple cleaning
```{r, message=FALSE}
# covid.confirmed <- covid.confirmed.raw %>%
#   select(-c(UID, iso2, iso3, code3, FIPS, Country_Region, Lat, Long_, Combined_Key)) %>%
#   rename(state = Province_State,
#          county = Admin2) %>%
#   group_by(county) %>% # Specify group indicator
#   # summarise(across(everything(), sum)) %>% # Specify column
#   pivot_longer(-c(state, county),
#                names_to = "date",
#                values_to = "cumulative.confirmed.cases") %>%
#   filter(state %in% c(state.name, "District of Columbia")) %>%
#   group_by(county) %>%
#   mutate(confirmed.cases = cumulative.confirmed.cases - lag(cumulative.confirmed.cases),
#          date = as.Date(date, format = "%m/%d/%y"),
#          confirmed.cases = replace_na(confirmed.cases, 0),
#          county = paste(county, "County"))
# 
# # fix negative cumulative cases
# while (any(covid.confirmed$confirmed.cases < 0)) {
#   covid.confirmed <- covid.confirmed %>%
#     mutate(cumulative.confirmed.cases = ifelse(cumulative.confirmed.cases < lag(cumulative.confirmed.cases),
#                                          lag(cumulative.confirmed.cases), # if smaller than previous
#                                          cumulative.confirmed.cases), # if not
#            cumulative.confirmed.cases = replace_na(cumulative.confirmed.cases, 0),
#            confirmed.cases = cumulative.confirmed.cases - lag(cumulative.confirmed.cases),
#            confirmed.cases = replace_na(confirmed.cases, 0))
# }
# 
# dim(covid.confirmed) # , 
# saveRDS(covid.confirmed, file = 'datasets/covid confirmed.Rds')
```

### moving average cleaning
```{r, message=FALSE}
# covid.confirmed <- covid.confirmed.raw %>%
#   select(-c(UID, iso2, iso3, code3, FIPS, Country_Region, Lat, Long_, Combined_Key)) %>%
#   rename(state = Province_State,
#          county = Admin2) %>%
#   group_by(county) %>% # Specify group indicator
#   # summarise(across(everything(), sum)) %>% # Specify column
#   pivot_longer(-c(state, county),
#                names_to = "date",
#                values_to = "cumulative.confirmed.cases") %>%
#   filter(state %in% c(state.name, "District of Columbia")) %>%
#   group_by(county) %>%
#   mutate(confirmed.cases = cumulative.confirmed.cases - lag(cumulative.confirmed.cases),
#          date = as.Date(date, format = "%m/%d/%y"),
#          confirmed.cases = replace_na(confirmed.cases, 0),
#          county = paste(county, "County"))
# 
# # fix negative cumulative cases
# covid.confirmed.imputed <- covid.confirmed %>%
#   group_by(county) %>%
#   # a lot of fluctuations so simple avg
#   mutate(cumulative.confirmed.cases = ifelse(cumulative.confirmed.cases < lag(cumulative.confirmed.cases),
#                                              NA, cumulative.confirmed.cases),
#          cumulative.confirmed.cases = na_ma(cumulative.confirmed.cases, weighting = "simple"),
#          confirmed.cases = cumulative.confirmed.cases - lag(cumulative.confirmed.cases))
# 
# # dim(covid.confirmed.imputed) # 1311362, 5
# saveRDS(covid.confirmed.imputed, file = 'datasets/covid confirmed moving average.Rds')
```

Just plotting something out~
```{r}
# covid.confirmed[1:5000,] %>%
#   ggplot(aes(x = date, y = cumulative.confirmed.cases, color = county)) +
#   geom_line() +
#   scale_color_hue(name='Midwest States')+
#   scale_x_date(date_breaks = "1 month", date_labels = "%b") +
#   # ylim(0, 1000000)+
#   ylab("Cumulative Confirmed Cases") +
#   xlab("Date") +
#   ggtitle("Cumulative Confirmed Cases in Midwest US") +
#   theme_minimal() +
#   theme(axis.title = element_text(size=12), # change axis style
#         legend.title = element_text(size=12), # change legend style
#         plot.title = element_text(size=14)) # change title style
```


## Cleaning Google mobility data

Importing Google mobility dataset
```{r, message=FALSE}
US_Region_2020_mobility <- read.csv( "./datasets/2020_US_Region_Mobility_Report.csv", header=T, na.strings=c("","NA"))

US_Region_2020_mobility <- US_Region_2020_mobility %>% 
  filter(sub_region_2 != "Null") %>% 
  rename(county = sub_region_2,
         state = sub_region_1)

# unique(US_State_2020_mobility_without_county$sub_region_1)
StateMonthMobility <- US_Region_2020_mobility %>%
  mutate(date = as.Date(date, "%Y-%m-%d"))
```

## combining covid and mobility dataset (simple cleaning)
```{r}
# covid.confirmed <- readRDS("datasets/covid confirmed.Rds")
# 
# length(unique(covid.confirmed$county)) # 1901
# length(unique(StateMonthMobility$county)) # 1718
# length(unique(svi.data$county)) # 1851
# 
# covid.mobility <- covid.confirmed %>%
#   left_join(StateMonthMobility, by = c("state", "county","date")) %>%
#   select(county, state, date, cumulative.confirmed.cases, confirmed.cases,
#          retail_and_recreation_percent_change_from_baseline,
#          grocery_and_pharmacy_percent_change_from_baseline,
#          parks_percent_change_from_baseline,
#          transit_stations_percent_change_from_baseline,
#          workplaces_percent_change_from_baseline) %>%
#   mutate(STATE = toupper(state))
# 
# covid.mob.svi <- covid.mobility %>%
#   left_join(svi.data, by = c("STATE", "county")) %>%
#   select(-c(STATE, COUNTY, FIPS)) %>%
#   rename(population = E_TOTPOP)
# 
# dim(covid.mob.svi) # 1311362, 11
# length(unique(covid.mob.svi$county)) # 1901
# saveRDS(covid.mob.svi, file = "datasets/covid mobility svi.Rds")
```

## combining covid and mobility dataset (moving average cleaning)
```{r}
# covid.confirmed <- readRDS("datasets/covid confirmed moving average.Rds")
# 
# length(unique(covid.confirmed$county)) # 1901
# length(unique(StateMonthMobility$county)) # 1718
# length(unique(svi.data$county)) # 1851
# 
# covid.mobility <- covid.confirmed %>%
#   left_join(StateMonthMobility, by = c("state", "county","date")) %>%
#   select(county, state, date, cumulative.confirmed.cases, confirmed.cases,
#          retail_and_recreation_percent_change_from_baseline,
#          grocery_and_pharmacy_percent_change_from_baseline,
#          parks_percent_change_from_baseline,
#          transit_stations_percent_change_from_baseline,
#          workplaces_percent_change_from_baseline) %>%
#   mutate(STATE = toupper(state))
# 
# covid.mob.svi <- covid.mobility %>%
#   left_join(svi.data, by = c("STATE", "county")) %>%
#   select(-c(STATE, COUNTY, FIPS)) %>%
#   rename(population = E_TOTPOP)
# 
# dim(covid.mob.svi) # 1311362, 11
# length(unique(covid.mob.svi$county)) # 1901
# saveRDS(covid.mob.svi, file = "datasets/covid mobility svi moving average.Rds")
```

## growth rate 
```{r, message=FALSE, warning=FALSE}
# covid.mob.svi <- readRDS('datasets/covid mobility svi.Rds')
covid.mob.svi <- readRDS('datasets/covid mobility svi moving average.Rds')
# dim(covid.mob.svi) # 1311362, 11

start_day <- "2020-02-01"
end_day <- "2020-12-31"
# active rates df
covid.mob.svi <- covid.mob.svi %>%
  mutate(# confirmed.cases = confirmed.cases + 1,
         # cumulative.confirmed.cases = cumulative.confirmed.cases + 1,
         confirmed.rates = 100 * confirmed.cases / population) %>%
  select(state, county, date, cumulative.confirmed.cases, confirmed.cases, population, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline, parks_percent_change_from_baseline, transit_stations_percent_change_from_baseline, workplaces_percent_change_from_baseline, confirmed.rates) %>%
  filter(date >= as.Date(start_day) & date <= as.Date(end_day)) %>%
  filter(!is.na(population)) 

# sum(is.na(covid.mob.svi$population))
# dim(covid.mob.svi) # 815826, 12

covid.time.series <- covid.mob.svi %>%
  mutate(county = paste(state, county)) %>%
  group_by(county) %>%
  mutate(avg.3days = (rollsumr(confirmed.rates, k = 3, fill = NA))/3 + 0.5,
         avg.7days = (rollsumr(confirmed.rates, k = 7, fill = NA))/7 + 0.5) 

lagged_days <- 11
covid.time.series <- covid.time.series %>%
  group_by(county) %>%
  mutate(growth.rate = log(avg.3days)/log(avg.7days)) %>% # other ways of calculating growth rate?
  # difference between the 2 numbers or the log of the ratio
  filter(date >= as.Date(start_day) + 6) %>%
  group_by(county) %>%
  mutate(growth.rate.lag = lag(growth.rate, lagged_days)) %>% # 2020-04-26 for one week lag
  filter(date >= as.Date(start_day) + 6 + lagged_days)
# 
# covid.time.series <- covid.time.series %>%
#   group_by(county) %>%
#   mutate(growth.rate = (cumulative.confirmed.cases - lag(cumulative.confirmed.cases)) / lag(cumulative.confirmed.cases)) %>% # other ways of calculating growth rate?
#   # difference between the 2 numbers or the log of the ratio
#   filter(date >= as.Date(start_day)) %>%
#   group_by(county) %>%
#   mutate(growth.rate.lag = lag(growth.rate, lagged_days)) %>% # 2020-04-26 for one week lag
#   filter(date >= as.Date(start_day) + lagged_days)

sum(is.na(covid.time.series$growth.rate)) # 0
covid.time.series[is.na(covid.time.series$growth.rate),]

# covid.time.series %>%
#   group_by(county) %>%
#   summarize(n = n())
#
# dates.gr <- format(seq(from = as.Date("2020/04/19"), to = as.Date("2021/02/15"), by = "day"), "%Y-%m-%d")
#
# for (date in dates.gr){
#   covid.time.series <- covid.time.series %>%
#   group_by(county) %>%
#   mutate(growth.rate = log(avg.3days)/log(avg.7days))
# }

dim(covid.time.series) # 763092, 16
saveRDS(covid.time.series, file = "datasets/covid time series.Rds")
```

### check fluctuations, plot to see which imputation to use (simple)
```{r}
# covid.time.series[1:1000,] %>%
#   ggplot(aes(x = date, y = confirmed.rates, color = county)) +
#   geom_line()
```


# Data analysis


## PCA 
```{r, message=FALSE}
# start running here~~~~~!!!!
covid.time.series <- readRDS("datasets/covid time series.Rds")
# dim(covid.time.series) # 985962, 12

# filter out small population counties 
covid.time.series <- covid.time.series %>%
  filter(population > 100000) 
range(covid.time.series$population)

length(c(unique(covid.time.series$county))) # 587

sum(is.na(covid.time.series$growth.rate)) # 0

# covid.time.series.1 <- covid.time.series %>% na.omit
covid.time.series.2 <- covid.time.series %>%
  select(-parks_percent_change_from_baseline)
# covid.time.series.3 <- covid.time.series.2 %>% na.omit
# l1 = unique(covid.time.series$county)
# l3 = unique(covid.time.series.3$county)

# covid.time.series.2 <- covid.time.series

imputation.counts <- covid.time.series.2 %>%
  group_by(county) %>% 
  summarize(count = n()) %>% 
  pull(count)

covid.time.series.nas <- covid.time.series.2 %>%
  group_by(county) %>% 
  summarise_each(funs(sum(is.na(.)))) %>%
  filter(retail_and_recreation_percent_change_from_baseline <= imputation.counts[1] - 3 & 
           # parks_percent_change_from_baseline <= imputation.counts[1] - 3 & 
           grocery_and_pharmacy_percent_change_from_baseline <= imputation.counts[1] - 3 & 
           transit_stations_percent_change_from_baseline <= imputation.counts[1] - 3 & 
           workplaces_percent_change_from_baseline <= imputation.counts[1] - 3)

county_nonnas = c(covid.time.series.nas$county)
length(county_nonnas) # 484
covid.time.series.nonnas <- covid.time.series.2 %>% 
  filter(county %in% county_nonnas)

sapply(covid.time.series.nonnas, function(x) sum(is.na(x)))
# data3 <- data[rowSums(is.na(data)) == 0, ]
# covid.time.series.nas <- covid.time.series.nonnas[, colSums(is.na(covid.time.series.nonnas)) <= 2]
```

### PCA imputation 
```{r}
covid.time.series.imputed <- covid.time.series.nonnas %>%
  group_by(county) %>%
  # a lot of fluctuations so simple avg
  mutate(retail_and_recreation_percent_change_from_baseline = 
           na_ma(retail_and_recreation_percent_change_from_baseline, weighting = "simple"),
         grocery_and_pharmacy_percent_change_from_baseline = 
           na_ma(grocery_and_pharmacy_percent_change_from_baseline, weighting = "simple"),
         # parks_percent_change_from_baseline = 
         #   na_ma(parks_percent_change_from_baseline, weighting = "simple"),
         transit_stations_percent_change_from_baseline = 
           na_ma(transit_stations_percent_change_from_baseline, weighting = "simple"),
         workplaces_percent_change_from_baseline = 
           na_ma(workplaces_percent_change_from_baseline, weighting = "simple"))

sapply(covid.time.series.imputed, function(x) sum(is.na(x)))

library(robustHD)
selected_cols <- c("retail_and_recreation_percent_change_from_baseline", "grocery_and_pharmacy_percent_change_from_baseline",
                   "transit_stations_percent_change_from_baseline", "workplaces_percent_change_from_baseline")
                   # "parks_percent_change_from_baseline")
covid.time.series.imputed[,selected_cols] = standardize(covid.time.series.imputed[,selected_cols])
covid.time.series.imputed
```

### PCA calculation 
PCA calculation
```{r}
# library(missMDA)
# filling missing data using iterative imputation
# nb <- estim_ncpPCA(mobility.active.rates[,3:7], method.cv = "Kfold", verbose = FALSE) # estimate # components
#(available methods include GCV to approximate CV)
# nb$ncp #4
# mobility.active.comp <- imputePCA(mobility.active.rates[,3:7], ncp = nb$ncp) # iterativePCA algorithm

covid.mobility.pca <- prcomp(covid.time.series.imputed[,selected_cols], center = TRUE, scale = TRUE)
covid.mobility.pca$rotation <- -covid.mobility.pca$rotation
covid.mobility.pca$x <- -covid.mobility.pca$x # negation necessary?
# biplot(mobility.pca, scale = 0)

# combine imputed data with growth rate 
# add PC1 as pca
covid.time.series.imputed$covid.mobility.pca = covid.mobility.pca$x[,"PC1"]
# equal weight 
glimpse(covid.time.series.imputed)
```

## biweekly binning 
```{r}
covid.final <- covid.time.series.imputed %>%
  mutate(month = months(date),
         week = format(date, format = '%W'),
         # the following code converts to biweekly
         biweek = 14 * (as.numeric(date - min(date)) %/% 14) + min(date)) %>%
  group_by(county, biweek) %>%
  summarise(confirmed.avg = mean(confirmed.rates, na.rm=TRUE),
            retail.recreation.avg = mean(retail_and_recreation_percent_change_from_baseline, na.rm=TRUE),
            grocery.pharmacy.avg = mean(grocery_and_pharmacy_percent_change_from_baseline, na.rm=TRUE),
            # parks.avg = mean(parks_percent_change_from_baseline, na.rm=TRUE),
            transit.avg = mean(transit_stations_percent_change_from_baseline, na.rm=TRUE),
            workplace.avg = mean(workplaces_percent_change_from_baseline, na.rm=TRUE),
            mobility.pca.avg = mean(covid.mobility.pca, na.rm = TRUE),
            growth.rate.avg = mean(growth.rate, na.rm = TRUE))

covid.small.county <- c(unique(covid.final[covid.final$growth.rate.avg > 1.05,]$county))

# exclude county with impossible large gr
covid.final <- covid.final %>%
  filter(!county %in% covid.small.county)

# covid.final[covid.final$growth.rate.avg > 1.5,]
# covid.final[covid.final$county== 'Texas Midland County',]
# covid.time.series[covid.time.series$county== 'Texas Midland County',]


covid.large.county <- covid.time.series.imputed %>% 
  arrange(desc(population)) %>%
  pull(county)
covid.large.county = c(unique(covid.large.county))
covid.large.county = covid.large.county[1:25]


covid.final %>% 
  # subset(!county %in% c("New York Suffolk County", "New York Queens County", "California Los Angeles County", "Pennsylvania Philadelphia County", "Texas Bextar County")) %>% 
  ggplot(aes(x = mobility.pca.avg, y = growth.rate.avg)) +
  geom_point(size = 0.2, alpha = 0.2) +
  geom_smooth(method = "lm", se  = FALSE) +
  theme(legend.position = 'none')
```

# growth rate & mobility rate by date
```{r}
#install package TTR()
library("TTR")
covid.time.series.imputed %>%
  filter((county %in% covid.large.county) & (!county %in% covid.small.county)) %>%
  # filter(date >= as.Date("2020-05-01") & date <= as.Date("2020-07-01")) %>% 
  ggplot(aes(x = date, y = growth.rate, color = county)) +
  geom_point(size = 0.5, alpha = 0.2) +
  geom_smooth(se  = FALSE) +
  theme(legend.position = 'none')

covid.time.series.imputed %>%
  filter((county %in% covid.large.county) & (!county %in% covid.small.county)) %>%
  # filter(date >= as.Date("2020-02-01") & date <= as.Date("2020-05-01")) %>% 
  ggplot(aes(color = county)) +
  geom_point(aes(x = date, y = covid.mobility.pca), size = 0.5, alpha = 0.2) +
  geom_smooth(aes(x = date, y = covid.mobility.pca), se  = FALSE, 
              formula = y ~ poly(x, 2)
              # span = 1
              ) +
  # geom_line(aes(x = date, y = SMA(covid.mobility.pca, n=20))) + 
  theme(legend.position = 'none')

covid.time.series.imputed %>%
  filter((county %in% covid.large.county) & (!county %in% covid.small.county)) %>% 
  # subset(!county %in% c("New York Suffolk County", "New York Queens County", "California Los Angeles County", "Pennsylvania Philadelphia County", "Texas Bextar County")) %>% 
  ggplot(aes(x = covid.mobility.pca, y = growth.rate, color = county)) +
  geom_point(size = 0.8, alpha = 0.2) +
  geom_smooth(method = "lm", se  = FALSE) +
  theme(legend.position = 'none')
```

## residual to residual analysis 
```{r}
library(lme4)
# add states to the lm() model?
covid.final <- covid.final %>%
  filter(county %in% covid.large.county)
mobility.active.country.lm <- lm(growth.rate.avg ~ mobility.pca.avg, data = covid.final)
summary(mobility.active.country.lm)

# lagged glm
mobility.active.country.lagged.lm <- lm(growth.rate.lag ~ covid.mobility.pca, 
                                        data = covid.time.series.imputed %>% 
                                          filter(county %in% covid.large.county & !county %in% covid.small.county))
summary(mobility.active.country.lagged.lm)

# lmer with county as within
mobility.active.country.lmer <- lmer(growth.rate.avg ~ mobility.pca.avg + (1|county), data = covid.final)
summary(mobility.active.country.lmer)

mobility.active.country.res <- residuals(mobility.active.country.lm)
# retail and recreation
retail.active.country.lm <- lm(retail.recreation.avg ~ mobility.pca.avg, data = covid.final)
retail.bottom.res <- residuals(retail.active.country.lm)
# below is lm(residual ~ residual)
retail.country.lm <- lm(mobility.active.country.res ~ retail.bottom.res)
retail.country.sum <- summary(retail.country.lm)
# r squared = 3.911e-05, p = 0.381


# grocery and pharmacy
grocery.pharmacy.mobility.country.lm <- lm(grocery.pharmacy.avg ~ mobility.pca.avg, data = covid.final)
grocery.pharmacy.mobility.country.res <- residuals(grocery.pharmacy.mobility.country.lm)
# below is lm(residual ~ residual)
grocery.country.lm <- lm(mobility.active.country.res ~ grocery.pharmacy.mobility.country.res)
grocery.country.sum <- summary(grocery.country.lm)
# r squared =   0.003586, p <2e-16 ***

# # parks
# parks.mobility.country.lm <- lm(parks.avg ~ mobility.pca.avg, data = covid.final)
# # below is lm(residual ~ residual)
# parks.mobility.country.res <- residuals(parks.mobility.country.lm)
# parks.country.lm <- lm(mobility.active.country.res ~ parks.mobility.country.res)
# parks.country.sum <- summary(parks.country.lm)
# # r squared = 0.07035

# transit
transit.mobility.country.lm <- lm(transit.avg ~ mobility.pca.avg, data = covid.final)
# below is lm(residual ~ residual)
transit.mobility.country.res <- residuals(transit.mobility.country.lm)
transit.country.lm <- lm(mobility.active.country.res ~ transit.mobility.country.res)
transit.country.sum <- summary(transit.country.lm)
# r squared = 9.925e-05, p = 0.163

# workplace
workplace.mobility.country.lm <- lm(workplace.avg ~ mobility.pca.avg, data = covid.final)
workplace.mobility.country.res <- residuals(workplace.mobility.country.lm)
# below is lm(residual ~ residual)
workplace.country.lm <- lm(mobility.active.country.res ~ workplace.mobility.country.res)
workplace.country.sum <- summary(workplace.country.lm)
# r squared =  0.0042, p<2e-16 ***

residual.lm <- lm(growth.rate.avg ~ mobility.pca.avg + retail.bottom.res + grocery.pharmacy.mobility.country.res + transit.mobility.country.res + workplace.mobility.country.res, data = covid.final)
summary(residual.lm)
```

```{r}
covid.final %>%
  ggplot(aes(x = mobility.pca.avg,y = growth.rate.avg)) + 
  geom_point() 



```






```{r}
final.state = sapply(strsplit(covid.final$county, ' '), function(x) x[1])
covid.final.1 <- covid.final 
covid.final.1$state = final.state
unique(covid.final.1$state)
covid.nh <- covid.final.1 %>%
  filter(state == "New")



```



```{r}
covid.ca <- covid.final.1 %>%
  filter(state == "California")

mobility.active.ca.lm <- lm(growth.rate.avg ~ mobility.pca.avg, data = covid.ca)
mobility.active.ca.res <- residuals(mobility.active.ca.lm)
# retail and recreation
retail.active.ca.lm <- lm(retail.recreation.avg ~ mobility.pca.avg, data = covid.ca)
retail.ca.res <- residuals(retail.active.ca.lm)
# below is lm(residual ~ residual)
retail.ca.lm <- lm(mobility.active.ca.res ~ retail.ca.res)
retail.ca.sum <- summary(retail.ca.lm)
# r squared = 0.001367, p = 0.262


# grocery and pharmacy
grocery.pharmacy.mobility.ca.lm <- lm(grocery.pharmacy.avg ~ mobility.pca.avg, data = covid.ca)
grocery.pharmacy.mobility.ca.res <- residuals(grocery.pharmacy.mobility.ca.lm)
# below is lm(residual ~ residual)
grocery.ca.lm <- lm(mobility.active.ca.res ~ grocery.pharmacy.mobility.ca.res)
grocery.ca.sum <- summary(grocery.ca.lm)
# r squared =   0.01534, p = 0.00016 ***

# parks
#parks.mobility.country.lm <- lm(parks.avg ~ mobility.pca.avg, data = covid.final)
# below is lm(residual ~ residual)
#parks.mobility.country.res <- residuals(parks.mobility.country.lm)
#parks.country.lm <- lm(mobility.active.country.res ~ parks.mobility.country.res)
#parks.country.sum <- summary(parks.country.lm)
# r squared = 0.07035

# transit
transit.mobility.ca.lm <- lm(transit.avg ~ mobility.pca.avg, data = covid.ca)
# below is lm(residual ~ residual)
transit.mobility.ca.res <- residuals(transit.mobility.ca.lm)
transit.ca.lm <- lm(mobility.active.ca.res ~ transit.mobility.ca.res)
transit.ca.sum <- summary(transit.ca.lm)
# r squared = 0.002228, p = 0.152

# workplace
workplace.mobility.ca.lm <- lm(workplace.avg ~ mobility.pca.avg, data = covid.ca)
workplace.mobility.ca.res <- residuals(workplace.mobility.ca.lm)
# below is lm(residual ~ residual)
workplace.ca.lm <- lm(mobility.active.ca.res ~ workplace.mobility.ca.res)
workplace.ca.sum <- summary(workplace.ca.lm)
# r squared =  0.002153, p= 0.159

residual.ca.lm <- lm(growth.rate.avg ~ mobility.pca.avg + retail.ca.res + grocery.pharmacy.mobility.ca.res + transit.mobility.ca.res + workplace.mobility.ca.res, data = covid.ca)
summary(residual.ca.lm)
summary(mobility.active.ca.lm)
# positive coefficient, r-suqred = 0.01447, p = 0.000247 ***

```


```{r}
covid.tx <- covid.final.1 %>%
  filter(state == "Texas")

mobility.active.tx.lm <- lm(growth.rate.avg ~ mobility.pca.avg, data = covid.tx)
mobility.active.tx.res <- residuals(mobility.active.tx.lm)
# retail and recreation
retail.active.tx.lm <- lm(retail.recreation.avg ~ mobility.pca.avg, data = covid.tx)
retail.tx.res <- residuals(retail.active.tx.lm)
# below is lm(residual ~ residual)
retail.tx.lm <- lm(mobility.active.tx.res ~ retail.bottom.res)
retail.tx.sum <- summary(retail.tx.lm)
# r squared = 4.913e-05, p = 0.784


# grocery and pharmacy
grocery.pharmacy.mobility.tx.lm <- lm(grocery.pharmacy.avg ~ mobility.pca.avg, data = covid.tx)
grocery.pharmacy.mobility.tx.res <- residuals(grocery.pharmacy.mobility.tx.lm)
# below is lm(residual ~ residual)
grocery.tx.lm <- lm(mobility.active.tx.res ~ grocery.pharmacy.mobility.tx.res)
grocery.tx.sum <- summary(grocery.tx.lm)
# r squared =  0.01775, p = 1.65e-07 ***

# parks
#parks.mobility.country.lm <- lm(parks.avg ~ mobility.pca.avg, data = covid.final)
# below is lm(residual ~ residual)
#parks.mobility.country.res <- residuals(parks.mobility.country.lm)
#parks.country.lm <- lm(mobility.active.country.res ~ parks.mobility.country.res)
#parks.country.sum <- summary(parks.country.lm)
# r squared = 0.07035

# transit
transit.mobility.tx.lm <- lm(transit.avg ~ mobility.pca.avg, data = covid.tx)
# below is lm(residual ~ residual)
transit.mobility.tx.res <- residuals(transit.mobility.tx.lm)
transit.tx.lm <- lm(mobility.active.tx.res ~ transit.mobility.tx.res)
transit.tx.sum <- summary(transit.tx.lm)
# r squared = 3.697e-05, p = 0.812

# workplace
workplace.mobility.tx.lm <- lm(workplace.avg ~ mobility.pca.avg, data = covid.tx)
workplace.mobility.tx.res <- residuals(workplace.mobility.tx.lm)
# below is lm(residual ~ residual)
workplace.tx.lm <- lm(mobility.active.tx.res ~ workplace.mobility.tx.res)
workplace.tx.sum <- summary(workplace.tx.lm)
# r squared =  0.03123, p = 3.2e-12 ***

residual.tx.lm <- lm(growth.rate.avg ~ mobility.pca.avg + retail.tx.res + grocery.pharmacy.mobility.tx.res + transit.mobility.tx.res + workplace.mobility.tx.res, data = covid.tx)
summary(residual.lm)
summary(mobility.active.tx.lm)
# negative coefficient, r-squared = 0.02968, p = 1.12e-11 ***

```

